---
title: "Skin (CXCR6) scRNA-seq Analysis"
author: "Austin Schultz (GitHub: aschultz37)"
date: "February 10, 2023"
output: html_document
---

# Initialization Code
Run this section once **at the beginning of each R session**.
```{r initialization, echo=TRUE, message=FALSE, results='hide'}
library(dplyr)
library(Seurat)
library(patchwork)
library(RColorBrewer)
library(EnhancedVolcano)

custom_color_palette <-  c("#1B9E77", "#D95F02", "#7570B3", "#CA00C4",
                          "#66A61E", "#E6AB02", "#A6761D", "#666666", "#194D33")

LN_Trm_genes <- c("Acap1", "Actn2", "Amica1", "Arhgef1", "Atxn7l3b", "Aw112010",
                  "B4galnt1", "Bcl11b", "Cbx3", "Ccnd2", "Ccr10", "Cd27", "Cd7",
                  "Cd74", "Chd3", "Cirbp", "Clec2d", "Crot", "Csf1", "Cxcr3",
                  "Cxcr6", "Eif5", "Evl", "Fam189b", "Fubp1", "Fyb", "Gramd1a",
                  "Sema4a","Gstp1", "Shisa5", "H2-T23", "Sipa1", "Hmgn1",
                  "Slfn2", "Hmha1", "Sp100", "Hsp90b1", "Spcs2", "Id2", "Srrm2",
                  "Ifitm10", "Stap1", "Ikzf3", "Tbc1d10c", "Il16", "Tesc",
                  "Il18r1", "Tnfaip8", "Il7r", "Tnrc6a", "Irf2bpl", "Tsc22d4",
                  "Itgae", "Uba52", "Itgal", "Ucp2", "Itm2c", "Wbp1", "Lfng",
                  "Xist", "Lpar6", "Ypel3", "Lrrc58", "Zbtb7a", "Ltb", "Znrf1",
                  "Ly6a", "Ly6e", "Ly6g5b", "Malat1", "Mbnl1", "Mrpl52", "Mxd4",
                  "Mycbp2", "N4bp2l2", "N4bp2l2", "Ndufa3", "Ndufa5", "Nktr",
                  "Nudcd3", "Ogt", "Pdcd4", "Pdia3", "Pdia6", "Ptpn7", "Ptprc",
                  "Rapgef6", "Rbpj", "Rgs10", "Rpl15", "Rpl35", "Rpl38",
                  "Rps28", "Rps29", "Sash3")

TGFb_genes <- c("Mcpt2","Cdh1","Spsb1","Vipr2","Gpr56","Src","Ppp2r2c","Lrig1",
                "Itgae","Agap1","Ncmap","Pmepa1","Sema6d","Emid1","Cd33","Dlk1",
                "Ldlrad4","Car2","Cpd","Nt5e","Tspan9","Gsg2","Klhl30",
                "1810011H11Rik","Osgin1","Ccl1","Litaf","Itga1","Kifc3",
                "Hsf2bp","Asic3","Abi3","Smurf2","Phactr2","Oplah","Qpct",
                "Tfr2","Isg20","Rnase6","Rgs1","2900026A02Rik","Mmp11",
                "Tnfsf11","Nrarp","Cyb561","Smyd1","Kcnip2","Cx3cr1","Nek6",
                "Nlrp1b","St8sia1","Arhgap39","Jup","Htra3","Rgs16","H2-M5",
                "Chn2","Cish","Atp6v0a1","Skil","Dok3","Igflr1","Ccr8","Timp2",
                "Zfyve28","Ppm1n","Hpgds","B4galnt4","Ifng","Ctnnal1","Clec12a",
                "Exoc3l","Coro2a","Ikzf4","Adamts6","D8Ertd82e","Smpd5","Aqp3",
                "Evpl","Ramp1","St8sia6","Xcl1","Scn1b","Rnf149","Dtx4","Gngt2",
                "Sbk1","Tbc1d16","Tnfrsf13c","Gna12","Ermn","Neu3","Fmnl3",
                "Cd83","Epb4.1l2","Ccdc112","Adam19","Rab26","Fam101b","Mical3",
                "Prkcz","Grina","Slc27a6","Tgfbr3","Fgfr1","Msc","Rgs10",
                "Lonrf1","Lax1","Kcnc1","Nphp1","Slc16a10","Kif13a","Ninj1",
                "Smyd3","9430020K01Rik","Csgalnact1","Gpaa1","Ski","Gcnt4",
                "Map9","Egr3","Fam161a","Egr1","Fndc3a","Mapkapk3","Ctss",
                "Hnrnpll","Galm","Dusp2","Stom","Esm1","1700049G17Rik",
                "Plekho1","Med10","Smtn","Gpr34","Sepn1","Egr2","Prrt2","Aen",
                "Cd101","Gtf2ird1","Tiam1","Camkk1","D430042O09Rik","Fam214b",
                "Matk","Ralgps1","Dapk2","Usp6nl","Foxred2","Wdyhv1","Znrf1",
                "Tjp1","Irf8","Hemk1","Pgap1","Accs","Aim2","Per3","Zfr2",
                "Lgalsl","1700001L05Rik","Zfp820","D3Ertd254e","Gcnt1",
                "Slc41a2","Ttc39b","Gclm","Peg13","Slc9a1","Adora3","Cers6",
                "Ccrn4l","Cd96","Golim4","Lpcat2","Lsr","Acsbg1","Eef2k",
                "Plekhf1","Rbm20","Ssx2ip","Ankrd50","Igfbp4","Inpp4b",
                "Irf2bpl","Pygl","Zfp1","Golm1","Gpr68","Ptgfrn","Tsc22d1",
                "Abca1","Fam124b","Itpripl2","Bcl6","Lysmd2","Trp53inp2",
                "Zdhhc13","Bpgm","F2r","Frmd4b","Ctsw","Swap70","Frmd6","Gas7",
                "Gdpd5","Spire1","Tet3","Batf3","Dstyk","Luzp1","Mgat5","Ptpre",
                "Ralgps2","Mif4gd","Stat1","Ttc3","Abhd15","Cerk","Adssl1",
                "Pcyt1b","Rai1","Blcap","Map3k14","Rnf19b","Scai","Tmem57",
                "Atp6v1g2","Chst12","Fam20a","Gtf3c1","Trp53inp1","Wdr78",
                "Aars2","Cd244","Ly6g5b","Tbx6","Usp22","Zfp827",
                "1600014C10Rik","Als2","Arhgef5","B4galt5","Nfat5","Prkacb",
                "Rgs2","Slc9a3r1","Soat1","Tctn3","Ttc39c","Cotl1","Ldlrap1",
                "Ncf1","Iigp1","Ikzf3","Ipcef1","Irf4","Abi2","Runx3","Ypel3",
                "Entpd1","Fut8","Inpp5f","Apol7e","Arhgef12","Nrp1","Slc26a11",
                "Tnfrsf1b","Cd160","Gfod1","Gm12185","H6pd","Pmm1","Tmem2",
                "Ublcp1","Dennd3","Gramd1a","Idh2","Ppip5k1","Slc39a13",
                "Baiap3","Extl3","Mxd1","Nipal1","Rrp1b","Twsg1","Cdc42bpg",
                "Celsr1","Ehd1","Kit","Slc22a15","Tmcc1","Camsap2","Klhl25",
                "Ncf4","Plcxd2","Rab11fip4","Specc1","Fam3c","Fuca2","Pde4a",
                "Prr12","Ctnnb1","Egln3","Fam46a","Fbxo25","Gprin3","Scly",
                "AW112010","Cd1d1","Lrrc61","Clstn1","Exosc4","Smad7","Susd3",
                "Traf4","Vasp","Gne","Gpbp1l1","Prkch","Rab37","Rbpj","Usp11",
                "AA467197","Bmpr2","Cd8a","Dpp9","Inpp5d","Kif1b","Lasp1",
                "Rftn1","Wee1","Fasl","Nbas","Plscr1","Prkdc","Rhoh","Spcs2",
                "Suox","Tbc1d4","Tgif1","Anp32a","Lnpep","Myo5a","Rreb1",
                "Zfp706","Ermp1","Fam149b","Glrx","Pacsin2","Plekha2","Sorl1",
                "Dnajc9","Nbeal1","Plod2","Ssh2","Trappc10","Ercc6","Fchsd2",
                "Gfi1","Ubn2","Vps54","Actr1b","Ccni","Cd2bp2","Tnfsf10",
                "Acot11","Atad2","Lgals9","Nup153","Gtpbp1")

TCR_genes <- c("Bcl10","Card11","Cd247","Cd3d","Cd3e","Cd3g","Cd4","Cdc34",
               "Chuk","Cul1","Fbxw11","Gm11808","H2aa","H2ab1","H2ea","H2eb1",
               "H2eb2","Ikbkb","Ikbkg","Inpp5d","Lck","Malt1","Map3k7","Nfkb1",
               "Nfkbia","Pdpk1","Pik3ca","Pik3cb","Pik3r1","Pik3r2","Prkcq",
               "Psma1","Psma2","Psma3","Psma4","Psma5","Psma6","Psma7","Psma8",
               "Psmb1","Psmb10","Psmb11","Psmb2","Psmb3","Psmb4","Psmb5",
               "Psmb6","Psmb7","Psmb8","Psmc1","Psmc2","Psmc3","Psmc4","Psmc5",
               "Psmc6","Psmd1","Psmd10","Psmd11","Psmd12","Psmd13","Psmd14",
               "Psmd2","Psmd3","Psmd4","Psmd5","Psmd6","Psmd7","Psmd8","Psmd9",
               "Psme1","Psme2","Psme2b","Psme3","Psme4","Psmf1","Pten","Rela",
               "Ripk2","Rps27a","Skp1","Tab2","Trac","Traf6","Trat1","Trav16",
               "Trav19","Trbc1","Trbc2","Trbv16","Uba52","Ubb","Ubc","Ube2d1",
               "Ube2d2a","Ube2n","Ube2v1")

TSCM_genes <- c("Slamf6", "Pdcd4", "Cd200", "Ltb", "Xcl1", "Irf3", "Cxcr5", 
                "Cd28", "Klf2", "Tox", "Tcf7", "Pdcd1")
```

# Initial QC, Pre-Processing, and Run
Create the Seurat Object.
```{r create_scobj}
scobj.data <- Read10X(data.dir="raw/")
scobj <- CreateSeuratObject(counts=scobj.data, project="2020-12-16_WT", 
                         min.cells=3, min.features=200)
```

## QC & Filtering
Calculate the % mitochondrial and ribosomal genes for each cell.
```{r qc_calc}
scobj[["percent.mt"]] <- PercentageFeatureSet(scobj, pattern="^mt-")
scobj[["percent.rps"]] <- PercentageFeatureSet(scobj, pattern="Rps")
scobj[["percent.rpl"]] <- PercentageFeatureSet(scobj, pattern="Rpl")
```

Visualize the QC features as violin plots.
```{r qc_vln}
VlnPlot(scobj, features=c("nFeature_RNA", "nCount_RNA"), ncol=2)
VlnPlot(scobj, features=c("percent.mt", "percent.rps", "percent.rpl"), ncol=3)
```

Visualize the QC features as FeatureScatter plots.
```{r qc_ftr}
plot1 <- FeatureScatter(scobj, feature1="nCount_RNA", feature2="percent.mt")
plot2 <- FeatureScatter(scobj, feature1="nCount_RNA", feature2="nFeature_RNA")
plot1 + plot2
```

Apply filters and normalize data.
```{r qc_filter}
scobj <- subset(scobj, subset=(nFeature_RNA > 100 & nFeature_RNA < 4000 
                & percent.mt < 5 & percent.rps < 20 & percent.rpl < 23))
scobj <- NormalizeData(scobj, normalization.method="LogNormalize",
                       scale.factor=10000)
```

## Identify Highly Variable Features
```{r find_hvg}
scobj <- FindVariableFeatures(scobj, selection.method="vst", nfeatures=2500)
```

## Scale Data
```{r scale_data}
all.genes <- rownames(scobj)
scobj <- ScaleData(scobj, features=all.genes)
```


## Linear Dimensional Reduction & Determine Dimensionality
Perform PCA.
```{r pca}
scobj <- RunPCA(scobj, features=VariableFeatures(object=scobj))
```

Jackstraw and Elbow Plots to visualize dimensionality of dataset.
```{r jackstraw}
scobj <- JackStraw(scobj, dims=30, num.replicate=100)
scobj <- ScoreJackStraw(scobj, dims=1:30)
JackStrawPlot(scobj, dims=1:30)
ElbowPlot(scobj, ndims=30)
```

## Cluster Cells
Run clustering and UMAP algorithms. Dims based on Jackstraw and Elbow plots.
```{r cluster}
scobj <- FindNeighbors(scobj, dims=1:17) 
scobj <- FindClusters(scobj, resolution=0.4)
scobj <- RunUMAP(scobj, dims=1:17)
```

Show the UMAP and the number of cells in each cluster.
```{r show_umap}
DimPlot(scobj, reduction="umap", pt.size=2.5, cols=custom_color_palette)
table(Idents(scobj))
```

## Generate Initial Heatmap
Find markers for all clusters.
```{r find_markers}
scobj.markers <- FindAllMarkers(scobj, only.pos=TRUE, min.pct=0.25,
                                logfc.threshold=0.25)
```

Generate heatmap for top 10 genes in all clusters.
```{r heatmap}
scobj.markers %>%
    group_by(cluster) %>%
    top_n(n=10, wt=avg_log2FC) -> top10
DoHeatmap(scobj, features=top10$gene, 
          group.colors=custom_color_palette) + NoLegend()
```

Make list of clusters we consider Trm and Tcirc. Excludes cluster 5 which
contains cycling cells.
```{r cluster_groups}
Trm_clusters = c(1, 2, 3)
Tcirc_clusters = c(0, 4)
```

Generate DEG list for Trm clusters vs T circ.
```{r Trm_DEG}
Trm_vs_Tcirc_markers <- FindMarkers(scobj, ident.1=Trm_clusters, 
                                    ident.2=Tcirc_clusters,
                                    min.pct=0.25)
write.csv(Trm_vs_Tcirc_markers, "output/Trm_vs_Tcirc_markers.csv", 
          row.names=TRUE)
```

LN Trm module score.
```{r Trm_modulescore}
scobj <- AddModuleScore(object=scobj, features=list(LN_Trm_genes),
                             ctrl=100, name="LN_Trm")
FeaturePlot(scobj, features="LN_Trm1", pt.size=2.5, 
            cols=rev(brewer.pal(n=11, name="RdBu")))
```

TGFb module score.
```{r tgfb_modulescore}
scobj <- AddModuleScore(object=scobj, features=list(TGFb_genes),
                             ctrl=100, name="TGFb")
FeaturePlot(scobj, features="TGFb1", pt.size=2.5, 
            cols=rev(brewer.pal(n=11, name="RdBu")))
```

TCR module score.
```{r tcr_modulescore}
scobj <- AddModuleScore(object=scobj, features=list(TCR_genes),
                             ctrl=100, name="TCR")
FeaturePlot(scobj, features="TCR1", pt.size=2.5, 
            cols=rev(brewer.pal(n=11, name="RdBu")))
```

TSCM module score.
```{r tscm_modulescore}
scobj <- AddModuleScore(object=scobj, features=list(TSCM_genes),
                             ctrl=100, name="TSCM")
FeaturePlot(scobj, features="TSCM1", pt.size=2.5, 
            cols=rev(brewer.pal(n=11, name="RdBu")))
```






